\documentclass[12pt]{../UWMadThesis}

\input{../HelperCommands}
\usepackage{enumitem}

%   Graphics path definition
\UWMadSetup{
    RelativeDirectory / {
        the-only-graphics-directory = Graphics
    }
}

%   Shim until merged back into the full document
\renewcommand{\Acronym}[1]{#1}

\ExplSyntaxOn

    \tl_set_eq:Nc \l_tmpa_tl {Gin@extensions}
    \clist_new:N \g__UWMad_RelativeDirectory_ImageExtensions_clist
    \clist_set:No
        \g__UWMad_RelativeDirectory_ImageExtensions_clist
        {\l_tmpa_tl}

    %\clist_show:N \g__UWMad_RelativeDirectory_ImageExtensions_clist
    \tl_new:N   \l__UWMad_File_Path_tl
    \tl_new:N   \l__UWMad_File_Name_tl
    \tl_new:N   \l__UWMad_File_Extension_tl
    
    \def\l_tmpa_tl{path/file.ext}
    
    \UWMad_File_PathFileName:nnnx
        \l__UWMad_File_Path_tl
        \l__UWMad_File_Name_tl
        \l__UWMad_File_Extension_tl
        {\l_tmpa_tl/\l_tmpa_tl}
        
    \tl_if_blank:VTF {\l__UWMad_File_Path_tl} {
        \msg_term:n {The~path~is~empty!}
    } {
        \msg_term:n {The~path~is~\l__UWMad_File_Path_tl}
    }
    \tl_show:N \l__UWMad_File_Name_tl
    \tl_show:N \l__UWMad_File_Extension_tl

\ExplSyntaxOff



\Institution{University of Wisconsin--Madison}


\begin{document}

\chapter{Introduction}

Stability of two-phase natural circulation systems is not a novel subject in and of itself.
However, this work aims to perform analysis on a novel geometry with unique characteristics using rigorous numerical methods.
Motivation for this effort is given through the examination of an experimental facility and preliminary data, both experimental and numerical.
Then, a literature overview of the field of stability analysis of fluid systems is presented with a broader perspective than that of this work.
Finally, the chapter concludes with the end goals of this investigation and an outline of the work.

\iffalse
 
\input{Section1-ReactorCavityCoolingSystem}
\newpage
\input{Section2-Literature}


\section{Research Purpose}\label{Section:Purpose}

The purpose of this research is to theoretically investigate the linear stability of a closed loop, natural circulation system undergoing phase-change with water as a working fluid.
The goal is to ultimately apply stability theory to three parallel channels in a closed circuit with mass loss, which form a novel combination for the literature.
As the system boils, the inventory is draining and a question that has not been completely answered is how does this affect the stability and overall evolution of the system.
In addition to the geometrical considerations, this work also explores a novel discretization scheme to the standard set of thermohydraulic equations that aims to more accurately represent branching flows than standard channel flow models.
Also, this work applies a full non-linear solver to the set of discretized equations with no operator splitting over a time-step.
Lastly, the linear stability of the system over a transient is determined through a detailed volume integration of the governing equations and an eigenvalue calculation.

The outline of the topics to be discussed in this work is:
\begin{enumerate}[topsep=0pt,parsep=0pt,itemsep=2pt]
    \item{conservation laws and pertinent thermohydraulic equations to solve;}
    \item{the discretization scheme and approximations applied to the equations;}
    \item{stability theory as applied to the set of nonlinear equations;}
    \item{the nonlinear solver used;}
    \item{results of transient simulations and associated stability calculations;}
    \item{conclusions from the presented results and future work.}
\end{enumerate}
Each section builds upon the last and works to fulfill the delineated goals above.



\bibliography{../Sources}
\bibliographystyle{unsrt}

 
 \fi

\end{document}




